---
title: "Getting Latitude and Longitude with Reutils"
author: "Amanda Alker"
date: "6/19/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Start with getting the right IDs and Search Queries


```{r}
#load the package
library(rentrez)
#This allows you to check out the databases and fields that you want to search
library(reutils)
#This is where the functions for getting db summaries and crosslinking the IDs are found
library(XML)
library(xml2)
library(rvest)
#These will help you parse the output
```


Check out the Searchable space...


```{r}
#Gives the list of searchable databases
entrez_dbs()
```



These are the searchable fields within the database of interest. If anything, perform the search with one example on the web interface to see where you need to be looking.



```{r}
#Gives the list of searchable terms within database 'sra'

entrez_db_searchable("sra")
```


Let's load in the test data.. 

```{r}
srr_list <- read.table("test.txt")
srr_list
str(srr_list)
```


Based on the Ids we have, it looks like we will have to find 'Biosample' Ids from the SRA run Ids. To do that, we will need to look up the uids for each database and convert them. 



#Convert the Ids


First, let's look for the uids that we want. This is just a practice with one.



```{r}
#make the 'uid' object to find the specific run ids in the sra database

uid <- esearch(srr_list$V1, db = "sra")
uid

#careful! These search queries are hard-coded, so we need to figure out a better way to have inputs for the function
```



So it looks like we got something out as a uid. Go and check it online to make sure you are getting the correct uid out..... it works!



```{r}
#find 'sra uid' from the SRA run id
sra_uid <- elink(uid, dbFrom = "sra", dbTo = "biosample" )
sra_uid
```



So now lets convert the sra_uid to the biosample_uid. Then let's double check that the right biosample_uid is coming up too. 




```{r}
#Reveal the biosample_uid that will be used to access the attribute metadata
biosample_uid <- linkset(sra_uid)
biosample_uid
```



Great! Now let's try to find the summary page of the biosample uid


#Find and fetch the data



```{r}
biosample_data <- esummary(biosample_uid, db = "biosample")
biosample_data
```



ITS IN THERE! This is the summary data, and in order to further parse out the "Attributes" section, we're going to need to do the full 'efetch'. 



##Parsing the data out of XML



```{r}
biosample_fetched <- efetch(biosample_uid, db = "biosample", retmode = "xml")
biosample_fetched
#this code did a good job of retreiving the xml data that could not be parsed out otherwise using different codes I tried. 

#parsed_biosample_fetched <- content(biosample_fetched, "parsed")
#The code above works, but not as well as the approach seen below.
```

We have the XML, now it's time to retreive the contents out. Unfortunately, whoever input the data for 'Attribute' didnt do a good job of discerning the important information. 


```{r}
attribute_contents <- biosample_fetched$xmlValue("//Attribute")
attribute_contents
#Here are the guts of the attribute section
```


Now that we have the values, we need to be able to get out the latitude and longitude only!


```{r}
attribute_df <- data.frame(attribute_contents)
attribute_df
#convert the output to a dataframe
```


```{r}
lat_lon <- attribute_df[seq(4, nrow(attribute_df), 5), ]
lat_lon
#Extract every 5th row, starting at the 4th position, going up to the number of rows found in the attribute_df
```



```{r}
lat_lon_df <- data.frame(lat_lon)
lat_lon_df
#Convert this to a data frame
```

```{r}
row.names(lat_lon_df) <- srr_list$V1
lat_lon_df
```



Wahoo! We've got it!



