---
title: "Class8 Computer Learning"
author: "Amanda Alker"
date: "4/27/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## K-means clustering example

Lets make up some data for testing the kmeans() function. 
```{r}
#two groupings of points, one centered at 3,-3
#cbind will turn our two datasets into columns and plot the temp and reverse temp 
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```


```{r}
km <- kmeans(x, centers = 2, nstart = 20)
km
#km gives list of 9 things, can specify which 
```

Use the kmeans() function setting k to 2 and nstart=20
Inspect/print the results

Q. How many points are in each cluster?
print km or will be in the global environment

Q. What ‘component’ of your result object details
 - cluster size?
 
 
```{r}
# km$size
```

 - cluster assignment/membership?

```{r}
# km$cluster
```
 
 - cluster center?

```{r}
# km$centers
```
 
Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
 
```{r}
plot(x, col= km$cluster, pch = 16)
points(km$centers, col = "blue", pch = 15)
```



##Hierarchical clustering

# First we need to calculate point (dis)similarity
# as the Euclidean distance between observations


```{r}
dist_matrix <- dist(x)
```

# The hclust() function returns a hierarchical
# clustering model
```{r}
hc <- hclust(d = dist_matrix)
# the print method is not so useful here- view only works as data.frames 
#pairwise distances 60x60 for how many objects 
hc 
```

```{r}
View(as.matrix(dist_matrix))
dim(as.matrix(dist_matrix))

```

convert to matrix to see the structure of the distance matrix and find the dimensions 

plot the clustering 

```{r}
plot(hc)
#split into two clusters and along the x
#lenght shows the degree of dissimilarity- more length = more disimilar
```

Use abline to draw the line at the height of where you want to split the tree


```{r}
plot(hc)
abline(h=6, col = "red")

```


Lets cut our tree to define our clusters 

```{r}
grps <- cutree(hc, k = 2)
table(grps)
```

```{r}
plot(x, col = grps )
```


# Step 1. Generate some example data for clustering
```{r}
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```

# Step 2. Plot the data without clustering

```{r}
plot(x)
```

# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
```{r}
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

```{r}
distmatrix <- dist(x)
```

```{r}
hc <- hclust(distmatrix)
plot(hc)
```

Lets cut into 3 groups 
```{r}
grps <- cutree(hc, k=3)
table(grps)
```

Plot the data colored by cluster
```{r}
plot(x, col = grps)
```


##Principal Component Analysis (PCA)

Import the data 
```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1)
head(mydata)
```

Lets do the PCA
```{r}
#need to transpose the data to run this analysis, because the function is old and they expect the rows and columns to be switched
#t(mydata)
pca <- prcomp(t(mydata), scale = TRUE)

#see what is returned by the prcomp() function
attributes(pca)
```

## A basic PC1 vs PC2 2-D plot

```{r}
#calling the columns [ ,#]
plot(pca$x[,1], pca$x[,2]) 
```

This provides the percentage that accounts for the variance or amount of data that is explained by the variable
```{r}
## Variance captured per PC
pca.var <- pca$sdev^2 
pca.var.per <- round(pca.var/sum(pca.var)*100, 1) 
pca.var.per
```


Now we create the barplot for the vector of numbers 
```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")

```

Now let's make this a little more useful and color the PCA plot

```{r}
## A vector of colors for wt and ko samples
#substr uses the x input, 1, 2 are the characters it wants to look at 
colvec <- as.factor( substr( colnames(mydata), 1, 2) )

plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 

#label all points with text() function
text(pca$x[,1], pca$x[,2], (labels = colnames(mydata)))
# OR
#identify(pca$x[,1], pca$x[,2], labels = colnames(mydata))
#allows you to click on things with a crosshair 
```



##Another Example

UK foods dataset

```{r}
#pay attention to the directory
x <- read.csv("/Users/mandialker/downloads/UK_foods.csv")
dim(x)
```

look at the header
```{r}
head(x)
```

change the rownames 
```{r}
rownames(x) <- x[,1]
#watch out for the numbers in the first column from last time. Shift the column names over by one 
x <- x[,-1]
head(x)
```

```{r}
dim(x)
```

```{r}
View(x)
knitr::kable(x, caption="The full UK foods data table")
```

Visualize with a heatmap
```{r}
#par(mar=c(20, 4, 4, 2))
heatmap(as.matrix(x))
```

Time for the PCA!!
```{r}
pca <- prcomp( t(x) )
summary(pca)
```

Plot the PCA and assign the y and y labels
```{r}
#xlim gives you the axis min and max
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```

Now let's calculate the percentages 
```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
```

```{r}
z <- summary(pca)
z$importance
```

This data can be summarized in a barplot 
```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```

```{r}
cumsum(v)
```

```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 4, 4, 2))
barplot( pca$rotation[,1], las=2 )
```


```{r}
## Or the inbuilt biplot() can be useful too 
biplot(pca)
```

