---
title: "Bioinformatics Class 12"
author: "Amanda Alker"
date: "5/11/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Call the package
```{r}
library(bio3d)
```

# Setup HIV-Pr for docking study

Get the protein
```{r}
file.name <- get.pdb("1hsg")
```

Read this file in and trim out the protein and small molecule ligand from everything else 

```{r}
hiv <- read.pdb(file.name)
#./ means that it lives locally
```

```{r}
ligand <- trim.pdb(hiv, "ligand")
ligand
```

Lets do the same for the protein
Extract protein

```{r}
protein <- trim.pdb(hiv, "protein")
protein
```

Write them up
```{r}
write.pdb(ligand, "1hsg_ligand.pdb")
write.pdb(protein, "1hsg_protein.pdb")
```


## Convert the docking results from pdbqt file to .pdb in 'Bio3d'

```{r}
res <- read.pdb("all.pdbqt", multi = TRUE)

write.pdb(res, "results.pdb")

```

Compare our results to Merk Co Crystal structure
```{r}
ori <- read.pdb("ligand.pdbqt")
```

```{r}
rmsd(ori, res)
```

>**Q6.** RMSD based on non hydrogen atoms 

```{r}
#Get protein (PDB, "protein")
inds <- atom.select(ori, "noh")

rmsd(ori$xyz[,inds$xyz], res$xyz[ , inds$xyz])
```

keep going in handout to check out the cool pdb capabilities, make a PCA, and superimpose the proteins on tope of each other 