---
title: "Bioinformatics Class7"
author: "Amanda Alker"
date: "4/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions again

We can source any file of R code with the 'source()' function
```{r}
source("http://tinyurl.com/rescale-R")
```

Lets make sure things are here

```{r}
ls()
```

check our 'rescale()' function is working

```{r}
rescale(1:10)
```

```{r, eval=FALSE}
rescale( c(1:10, "string") )
```

Let's check if 'rescale2()' does any better
```{r, eval=FALSE}
rescale2( c(1:10, "string") )
```

## Function for finding missing values in two datasets

Write a 'both_na() function to do this
```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

is.na(x)


```

provides where the NAs are located 
```{r}
which(is.na(x))
```

provides how many NAs there are 
```{r}
sum( is.na(x))
```

This will print these vectors seperately
```{r}
is.na(x)
is.na(y)
```

we can use & to combine the vectors
```{r}
sum(is.na(x) & is.na(y))
```

now we can make our function
```{r}
both_na <-function(x, y) {
  sum( is.na(x) & is.na(y))
}
```

test it
```{r}
both_na(x, y)
```

```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

both_na(x, y1)
both_na(x, y2)
```

```{r}
both_na2(x, y2)
```

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

ans <-both_na3(x, y)
```

```{r}
ans$which
```

##One more example- actually useful
```{r}
x <- df1$IDs
y <- df2$IDs

x
y
```

We can try the 'intersect()' function and the '%in%' functions

```{r}
intersect(x, y)
x %in% y
y %in% x

```

we can use the logical output in '%in% to get at our matching data..

```{r}
x[x %in% y]
y[y %in% x]
```


Lets put these together as columns of a matrix

```{r}
cbind(x[x %in% y], y[y %in% x])
```

```{r}
cbind( c("hello", "Help"), c("Please", "Me"))
rbind( c("hello", "Help"), c("Please", "Me"))
```

Now we can make our first function for this
```{r}
gene_intersect <- function(x, y) {
cbind(x[x %in% y], y[y %in% x])
}
```

Test it on x and y
```{r}
gene_intersect(x, y)
```

Lets change to take input data frames rather than vectors
```{r}
gene_intersect2(df1, df2) 
```

Looks good, this is our skateboard!

Add some flexibility for col name to match by... 
```{r}
gene_intersect3(df1, df2)
```

Make it look nicer- assign variables to the outputs- gene_intersect 4

Lets use the merge()function for this...
```{r}
merge(df1, df2, by="IDs")
```


